---
title: "hw6"
author: "Brennan Baker"
date: "November 26, 2018"
output:
  github_document:
    toc: true
---
# Problem 1

```{r load packages}
library(tidyverse)
```

### Load and tidy data

```{r load and tidy data}
homicides_df = read_csv(file = "./data-homicides-master/homicide-data.csv")

homicides_df = homicides_df %>% 
  mutate(city_state = str_c(city, ",\ ", state)) %>% 
  filter(!city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")) %>% 
  mutate(victim_age = as.numeric(victim_age)) %>% 
  filter(!is.na(victim_age)) %>% 
  filter(victim_race != "unknown") %>% 
  mutate(victim_race = ifelse(victim_race == "White", "white", "non-white")) %>% 
  mutate(victim_race = as.factor(victim_race)) %>% 
  mutate(victim_race = relevel(victim_race, ref = "white")) %>% 
  mutate(resolution = factor(ifelse(
    disposition %in% c("Closed without arrest", "Open/No arrest"), "unresolved", "resolved")))
```

After importing the data, I created a city_state variable. I excluded cities that have errors or do not report race. I converted age to numeric and removed NA age rows. I filtered out unknown race and recoded race as white or non-white, with white as the reference group. I removed unknown race because if the race is unknown, we cannot know if race is white or non-white. I recoded the disposition variavle as solved or unsolved. 

```{r baltimore fit}
baltimore_fit = homicides_df %>%
  filter(city_state == "Baltimore, MD") %>% 
  glm(resolution ~ victim_age + victim_race + victim_sex, data = ., family = binomial())
```

The above code saves a logistic regression to baltimore_glm. The model uses victim age, race, and sex to predict whether the case is resolved or unresolved.

```{r odds ratio}
baltimore_fit %>% broom::tidy(conf.int = TRUE) %>% 
  mutate(OR = exp(estimate),
         conf.low = exp(conf.low),
         conf.high = exp(conf.high)) %>% 
  select(term, OR, conf.low, conf.high) %>% 
  filter(term == "victim_racenon-white") %>% 
  knitr::kable(digits = 3)
```

The above table shows the estimate and confidence interval of the odds ratio for resolved vs unresolved homicides predicted by race, controling for age and sex.

```{r all cities fit, warning = FALSE}
homicides_models = homicides_df %>% 
  group_by(city_state) %>% 
  nest() %>% # now we have a list col called data containing all the data for each city state
  mutate(model = map(data, ~broom::tidy(glm(resolution ~ victim_age + victim_race + victim_sex, data = ., family = binomial()), conf.int = TRUE))) %>%
  select(-data) %>% 
  unnest() %>%
  mutate(OR = exp(estimate),
         conf.low = exp(conf.low),
         conf.high = exp(conf.high)) %>% 
  select(city_state, term, OR, conf.low, conf.high) %>%
  filter(term == "victim_racenon-white")

homicides_models %>% knitr::kable(digits = 3)
```

The above code creates runs the model that we used above for baltimore on each city_state.
